---
title: "Reproducible Research: Peer Assessment 1"
output: 
  html_document:
    keep_md: true
---


## Loading and preprocessing the data
### Load Libraries
```{r libraries}
library(data.table)
library(lubridate)
library(dplyr)
library(stringr)

```  

```{r read the data}
activity_raw <- read.csv("activity.csv")  
```
### See how it looks
```{r glance1}
str(activity_raw)
dt=data.table(activity_raw)
dt
```
### And a little bit more
```{r glance2}
summary(activity_raw)
```
## What is mean total number of steps taken per day?
```{r getdailysummary}
daily_act_raw <- tapply(activity_raw$steps, activity_raw$date, FUN=sum)
```

```{r showfreqs}
hist(daily_act_raw, breaks = 20, freq = NULL)
```  

### Change inline output so that numbers arent in scientific notation
```{r changeoptions} 
options(scipen = 1, digits = 2)
```
### mean daily steps
```{r getmean}
mean_steps <- mean(daily_act_raw, na.rm = TRUE)
```
The mean number of steps each day is `r round(mean_steps,0)`.   

### median daily steps
```{r getmedian}
med_steps <- median(daily_act_raw, na.rm = TRUE)
```
The median number of steps each day is `r med_steps`.   

## What is the average daily activity pattern?
Add a variable that makes the hours show up correctly in the plot and make a new dataset that we can manipulate without affecting the source dataset
```{r newdataset}
actdata <- activity_raw
actdata$newinterval <- ymd_hm(paste("2000-01-01", str_pad(activity_raw$interval, 4, pad = "0")))
```
And then take that data and create a file that groups it by interval
```{r groupbyinterval}
interval_grps <- group_by(actdata, newinterval, interval)
summry <- data.frame(summarise(interval_grps, mean(steps, na.rm=TRUE)))
```
Change the name of the resultant step summary field to make it easier to manipulate
```{r cleanupsummary}
summry$meansteps <- summry$mean.steps..na.rm...TRUE.
```
Plot that to see what is going on  
```{r plotdailyactivity}
plot(summry$newinterval, summry$meansteps, type="l")
```  
And the interval with the maximum number of average steps  
```{r maxstepsinterval}
maxstepint <- summry$interval[which.max(summry$meansteps)]
```  
The interval with the most steps, on average, is interval number `r maxstepint`.    

## Imputing missing values
1. Calculate and report the total number of missing values in the dataset (i.e. the total number of rows with `NA`s)
```{r calcnumnas}
numnas <- sum(is.na(activity_raw$steps))
```
There are `r numnas` NAs in the dataset.   

2. Devise a strategy for filling in all of the missing values in the dataset. The strategy does not need to be sophisticated. For example, you could use the mean/median for that day, or the mean for that 5-minute interval, etc.

My strategy is to take the average number of daily steps for each interval and apply those numbers to any intervals with missing values.  So, if interval 835 (the interval with the maximum average number of steps) were to have an NA in it for a particular day, that NA would be replaced by the 216.

3. Create a new dataset that is equal to the original dataset but with the missing data filled in.
#create a new dataset to house the imputed values
```{r imputemissings}
actDataImputed <- activity_raw
i <- 1
while(i < 17569) {
  if(is.na(actDataImputed$steps[i])) {
	actDataImputed$steps[i] <- summry$meansteps[which(actDataImputed$interval[i]==summry$interval)]
	}
	i <- i+1
}
```

```{r getdailyimputedsummary}
dayactDataImputed <- tapply(actDataImputed$steps, actDataImputed$date, FUN=sum)
```

4. Make a histogram of the total number of steps taken each day and Calculate and report the **mean** and **median** total number of steps taken per day. Do these values differ from the estimates from the first part of the assignment? What is the impact of imputing missing data on the estimates of the total daily number of steps?  

```{r histimputed}
hist(dayactDataImputed, breaks = 20, freq = NULL)
```  
### mean daily steps
```{r impmean}
imp_mean_steps <- mean(dayactDataImputed, na.rm = TRUE)
```
The mean number of steps each day is `r round(imp_mean_steps,0)`.   

### median daily steps
```{r impmedian}
imp_med_steps <- median(dayactDataImputed, na.rm = TRUE)
```
The median number of steps each day is `r imp_med_steps`.  
## Are there differences in activity patterns between weekdays and weekends?


